### 2013-10-31 index.Rmd
Changing the data munging into vector algebra

html-ouput                                                    source code
-------------------------------------------------------       -----------------------------
[loading and cleaning](code/descriptive_analysis.html)        [source](code/descriptive_analysis.Rmd)
[missing data & diagnostics](code/descriptive_analysis.html)  [source](code/descriptive_analysis.Rmd)

>>>>

html-ouput                                                    source code
-------------------------------------------------------       -----------------------------
[loading and cleaning](loadClean.html)                        [source](loadClean.Rmd)
[missing data & diagnostics](code/descriptive_analysis.html)  [source](code/descriptive_analysis.Rmd)

### 2013-10-31 loadClean.Rmd
Changing the data munging into vector algebra

```{rLoadClean1}

##################################################################
##################################################################
##################################################################
# Life in transition survey 2
# load and clean data

# libraries
library(car)
library(stringr)
library(foreign)

# LITS 2
# from web
#temp <- tempfile()
#download.file("http://www.ebrd.com/downloads/research/surveys/lits2.dta", temp)
#lits2 <- read.dta(temp)
# tai

load("~/workspace/data/lits/lits2.RData")
##################################################################
##################################################################

@@

```{rLoadClean2}
##################################################################
# different country groupings

# remove whitespaces from country
lits2$cntry <- str_trim(lits2$country_, side="right")
lits2$cntry <- factor(lits2$cntry)

lits2$socialism <- recode(lits2$cntry,
                            "c('France','Germany','Great Britain',
                            'Italy','Sweden')='Market economy';
                            c('Czech Republic','Estonia','Hungary',
                          'Bulgaria','Latvia','Lithuania','Poland',
                          'Slovakia','Slovenia','Romania','Albania',
                          'Bosnia','Bulgaria','Croatia','Macedonia',
                          'Montenegro','Kosovo','Serbia','Armenia',
                          'Azerbaijan','Belarus','Georgia','Kazakhstan',
                          'Kyrgyzstan','Moldova','Mongolia','Russia',
                          'Tajikistan','Ukraine','Uzbekistan')='Post-socialist';
                          else=NA")
lits2$group_analysis <- recode(lits2$cntry,
                             "c('France','Germany','Great Britain',
                             'Italy','Sweden')='Western Europe';
                             c('Czech Republic','Estonia','Hungary','Bulgaria',
                             'Latvia','Lithuania','Poland',
                             'Slovakia','Slovenia','Romania')='CEE';
                             c('Armenia','Azerbaijan','Belarus','Georgia','Kazakhstan',
                             'Kyrgyzstan','Moldova','Tajikistan',
                             'Ukraine','Uzbekistan','Russia')='CIS';else=NA")

@@

##

```{rLoadClean3}
################################
# reason for poverty
################################

################################
lits2$q309 <- as.factor(str_replace(lits2$q309, "Don't know", "Dont know"))

# All options

lits2$q309 <- recode(lits2$q309, "'Because of injustice in our society'='Social Blame';
                         'Because of laziness and lack of willpower'='Individual Blame';
                         'Because they have been unlucky'='Individual Fate';
                         'It is an inevitable part of modern life'='Social Fate';
                         'Not stated'='Not stated';
                         'Dont know'='Dont know'")

lits2$q309 <- factor(lits2$q309, levels=c("Social Blame","Individual Blame",
                                                  "Individual Fate","Social Fate",
                                          "Not stated","Dont know"))

# not stated and dont know coded NA

lits2$q309_rec <- recode(lits2$q309, "'Because of injustice in our society'='Social Blame';
                         'Because of laziness and lack of willpower'='Individual Blame';
                         'Because they have been unlucky'='Individual Fate';
                         'It is an inevitable part of modern life'='Social Fate';
                         c('Not stated','Dont know')=NA")

lits2$q309_rec <- factor(lits2$q309_rec, levels=c("Social Blame","Individual Blame",
                                                  "Individual Fate","Social Fate"))

################################
# $poverty - muuttuja
################################
#lits2$poverty <- recode(lits2$q309_rec, "'It is an inevitable part of modern life'=NA")
lits2$poverty <- lits2$q309_rec

################################
# $pov.log.social
################################
lits2$pov.log.social <- recode(lits2$q309_rec, "'Social Blame'=1;
                               else=0")
################################
# $pov_log_individual
################################
lits2$pov.log.individual <- recode(lits2$q309_rec, "'Individual Blame'=1;
                               else=0")

@@

##

```{rLoadClean4}

################################
# $sex - muuttuja
################################
lits2$sex <- lits2$q102_1

################################
# $income
################################

lits2$income <- factor(recode(lits2$q227, 
                           "1:3='Low income'; 
                          4:7='Middle income'; 
                           8:10='High income'; 
                           else=NA"))

lits2$income <- factor(lits2$income, levels=c("Low income","Middle income","High income"))

@@

# median(lits2)
#svyquantile(~q227, d.df, quantile = c(0.25, 0.5, 0.75), ci = TRUE)

lits2$q227_rec <- recode(lits2$q227, "-111:0=NA")
################################
# $future
################################
lits2$future <- factor(recode(lits2$q229, 
                              "1:3='a) low income'; 
                          4:7='b) middle income'; 
                              8:10='c) high income'; 
                              else=NA"))

@@

################################
# $past
################################
lits2$past <- factor(recode(lits2$q228, 
                              "1:3='a) low income'; 
                          4:7='b) middle income'; 
                              8:10='c) high income'; 
                              else=NA"))

@@

################################
# $age
################################

@@

################################
# $incsour tulonlähde
################################
lits2$q226m_txt <- recode(lits2$q226m, 
                          "1='Salary or wages in cash or in kind';
                                2='Income from self-employment';
                          3='Sales or bartering of farm products';
                          4='Pensions';5='Benefits from the state';
                          6='Help from relatives or friends';7='Other'")

@@

################################
# $edu
################################

@@

lits2$crise <- factor(lits2$crise, levels=c("Little or not at all",
                                            "Great or fair amount"))
summary(lits2$crise)
################################################################

@@

```{rLoadClean5}

################################################################
## Data for the analysis
df <- subset(lits2, select=c("SerialID","weight","cntry","country0","q309",
                             "q309_rec","sex",
                             "ageclass","age",
                             "future.diff","past.diff",
                             "pov.log.social",
                             "pov.log.individual",
                             "socialism","incsour3",
                             "edu2",
                             "group_analysis",
                             #"group_general","group_fine",
                             "income2",
                             "crise",
                             "q227", # income
                             "q228", # menneisyys
                             "q229", # tulevaisuus
                             "q226m_txt", # tulonlähde
                             "q515", # edu
                             "q801" # crise,
                             ))
# filter out not wanted countries
df <- df[!is.na(df$group_analysis),] # valitaan vaan group1 -muuttujan maat
df$cntry <- factor(df$cntry)
# print data in csv-format

################################################
################################################

@@

```{rLoadClean6}
################################################

@@

```{rLoadClean7}
#######
# Quality of governance
library(rustfare)
datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
                   'Ukraine','Moldova','Belarus','Kyrgyzstan',
                   'Kazakhstan','Tajikistan',
                   'Uzbekistan','Estonia','Latvia','Lithuania',
                   'Poland','Czech Republic','Slovakia','Hungary',
                   'Slovenia','Romania','Bulgaria',
                   'Italy','Germany','United Kingdom',
                   'Sweden','France'), 
       indicator = c("ffp_fsi", # Failed States Index
                     "wbgi_vae", # World Bank - Voice and Accountability
                     "undp_hdi", # UNDP Human Development Index
                     "fh_polity2", # Democracy (Freedom House/Polity)
                     "uw_gini"))
# subset the latest value per cntry per indicator
library(plyr)
datQogSum <- ddply(datQog, c("cname","indicator"), subset, year == max(year))

@@

################################

@@

```{rLoadClean8}
## GDP change from WB
### We need iso2c countrycode
library(countrycode)
cntry <- levels(df$cntry) # countrynames
iso2c <- countrycode(cntry, "country.name", "iso2c")
iso2c # copy paste into WDI search
library(WDI)
# WDIsearch(string='gdp', field='name', cache=NULL)
wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), country = c("AL","AM","AZ","BY","BA","BG","HR",
                                                       "CZ","EE","FR","GE","DE","GB","HU",
                                                       "IT","KZ","KG","LV","LT","MK","MD",
                                                       "MN","ME","PL","RO","RU","RS","SK",
                                                       "SI","SE","TJ","TR","UA","UZ"), 
           start = 2007, end = 2010)

>>>>

```{rLoadClean1}
# FROM WEB
#temp <- tempfile()
#download.file("http://www.ebrd.com/downloads/research/surveys/lits2.dta", temp)
#library(foreign)
#lits2 <- read.dta(temp)

# FROM LOCAL DISK
load("~/workspace/data/lits/lits2.RData")
```

@@

```{rLoadClean2}
# different country groupings
## remove whitespaces from country
library(stringr)
lits2$cntry <- str_trim(lits2$country_, side="right")
lits2$cntry <- factor(lits2$cntry)

lits2$socialism[lits2$cntry %in% c('France','Germany','Great Britain',
                                   'Italy','Sweden')]                   <- 'Market economy'

lits2$socialism[lits2$cntry %in% c('Czech Republic','Estonia','Hungary',
                                   'Bulgaria','Latvia','Lithuania',
                                   'Poland','Slovakia','Slovenia',
                                   'Romania','Albania','Bosnia',
                                   'Bulgaria','Croatia','Macedonia',
                                   'Montenegro','Kosovo','Serbia',
                                   'Armenia','Azerbaijan','Belarus',
                                   'Georgia','Kazakhstan','Kyrgyzstan',
                                   'Moldova','Mongolia','Russia',
                                   'Tajikistan','Ukraine','Uzbekistan')] <- 'Post-socialist'

summary(factor(lits2$socialism))

lits2$group_analysis[lits2$cntry %in% c('France','Germany',
                                        'Great Britain',
                                        'Italy','Sweden')] <- 'Western Europe'

lits2$group_analysis[lits2$cntry %in% c('Czech Republic','Estonia',
                                        'Hungary','Bulgaria',
                                        'Latvia','Lithuania',
                                        'Poland','Slovakia',
                                        'Slovenia','Romania')] <- 'CEE'

lits2$group_analysis[lits2$cntry %in% c('Armenia','Azerbaijan',
                                        'Belarus','Georgia',
                                        'Kazakhstan','Kyrgyzstan',
                                        'Moldova','Tajikistan',
                                        'Ukraine','Uzbekistan','Russia')] <- 'CIS'

summary(factor(lits2$group_analysis))

@@

## Recoding the dependent variable

```{rLoadClean3}
summary(lits2$q309)
# reason for poverty
library(stringr)
lits2$q309 <- as.factor(str_replace(lits2$q309, 
                                    "Don't know", 
                                    "Dont know"))
# All options
lits2$q309 <- as.character(lits2$q309)
lits2$q309[lits2$q309 %in% 'Because of injustice in our society'] <- 'Social Blame'
lits2$q309[lits2$q309 %in% 'Because of laziness and lack of willpower'] <- 'Individual Blame'
lits2$q309[lits2$q309 %in% 'Because they have been unlucky'] <- 'Individual Fate'
lits2$q309[lits2$q309 %in% 'It is an inevitable part of modern life'] <- 'Social Fate'
lits2$q309[lits2$q309 %in% 'Not stated'] <- 'Not stated'
lits2$q309[lits2$q309 %in% 'Dont know'] <- 'Dont know'

lits2$q309 <- factor(lits2$q309, levels=c("Social Blame",
                                          "Individual Blame",
                                          "Individual Fate",
                                          "Social Fate",
                                          "Not stated",
                                          "Dont know"))

summary(lits2$q309)
# not stated and dont know coded NA

lits2$poverty <- lits2$q309
lits2$poverty[lits2$poverty %in% c('Not stated','Dont know')] <- NA
lits2$poverty <- factor(lits2$poverty)
summary(lits2$poverty)

## For logistic regressions
# $pov.log.social

lits2$pov.log.social[lits2$q309 == 'Social Blame'] <- 1
lits2$pov.log.social[lits2$q309 != 'Social Blame'] <- 0

summary(factor(lits2$pov.log.social))

# $pov_log_individual
###

lits2$pov.log.individual[lits2$q309 == 'Individual Blame'] <- 1
lits2$pov.log.individual[lits2$q309 != 'Individual Blame'] <- 0

summary(factor(lits2$pov.log.individual))

@@

## Independent variables

```{rLoadClean4}
# $sex
lits2$sex <- lits2$q102_1
# $income

lits2$income[lits2$q227 == 1:3]  <- 'Low income'
lits2$income[lits2$q227 == 4:7]  <- 'Middle income'
lits2$income[lits2$q227 == 8:10] <-'High income'

lits2$income <- factor(lits2$income, 
                       levels=c("Low income",
                                "Middle income",
                                "High income"))

@@

# median(lits2)
#svyquantile(~q227, d.df, quantile = c(0.25, 0.5, 0.75), ci = TRUE)
library(car)
lits2$q227_rec <- recode(lits2$q227, "-111:0=NA")
###

# $future
###

lits2$future <- factor(recode(lits2$q229, 
                              "1:3='a) low income'; 
                          4:7='b) middle income'; 
                              8:10='c) high income'; 
                              else=NA"))

@@

###

# $past
###

lits2$past <- factor(recode(lits2$q228, 
                              "1:3='a) low income'; 
                          4:7='b) middle income'; 
                              8:10='c) high income'; 
                              else=NA"))

@@

###

# $age
###

@@

###

# $incsour tulonlähde
###

lits2$q226m_txt <- recode(lits2$q226m, 
                          "1='Salary or wages in cash or in kind';
                                2='Income from self-employment';
                          3='Sales or bartering of farm products';
                          4='Pensions';5='Benefits from the state';
                          6='Help from relatives or friends';7='Other'")

@@

###

# $edu
###

@@

lits2$crise <- factor(lits2$crise, levels=c("Little or not at all",
                                            "Great or fair amount"))
summary(lits2$crise)
###

@@

```{rLoadClean5, eval=TRUE}

###

## Data for the analysis
df <- subset(lits2, select=c("SerialID","weight","cntry","country0","q309",
                             "q309_rec","sex",
                             "ageclass","age",
                             "future.diff","past.diff",
                             "pov.log.social",
                             "pov.log.individual",
                             "socialism","incsour3",
                             "edu2",
                             "group_analysis",
                             #"group_general","group_fine",
                             "income2",
                             "crise",
                             "q227", # income
                             "q228", # menneisyys
                             "q229", # tulevaisuus
                             "q226m_txt", # tulonlähde
                             "q515", # edu
                             "q801" # crise,
                             ))
# filter out not wanted countries
df <- df[!is.na(df$group_analysis),] # valitaan vaan group1 -muuttujan maat
df$cntry <- factor(df$cntry)
# print data in csv-format

summary(df)

@@

```{rLoadClean6, eval=TRUE}
###

@@

```{rLoadClean7, eval=TRUE}
#######
# Quality of governance
library(rustfare)
datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
                   'Ukraine','Moldova','Belarus','Kyrgyzstan',
                   'Kazakhstan','Tajikistan',
                   'Uzbekistan','Estonia','Latvia','Lithuania',
                   'Poland','Czech Republic','Slovakia','Hungary',
                   'Slovenia','Romania','Bulgaria',
                   'Italy','Germany','United Kingdom',
                   'Sweden','France'), 
       indicator = c("ffp_fsi", # Failed States Index
                     "wbgi_vae", # World Bank - Voice and Accountability
                     "undp_hdi", # UNDP Human Development Index
                     "fh_polity2", # Democracy (Freedom House/Polity)
                     "uw_gini"))
# subset the latest value per cntry per indicator
library(plyr)
datQogSum <- ddply(datQog, c("cname","indicator"), subset, year == max(year))

@@

###

@@

```{rLoadClean8, eval=TRUE}
## GDP change from WB
### We need iso2c countrycode
library(countrycode)
cntry <- levels(df$cntry) # countrynames
iso2c <- countrycode(cntry, "country.name", "iso2c")
iso2c # copy paste into WDI search
library(WDI)
# WDIsearch(string='gdp', field='name', cache=NULL)
wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), country = c("AL","AM","AZ","BY","BA","BG","HR",
                                                       "CZ","EE","FR","GE","DE","GB","HU",
                                                       "IT","KZ","KG","LV","LT","MK","MD",
                                                       "MN","ME","PL","RO","RU","RS","SK",
                                                       "SI","SE","TJ","TR","UA","UZ"), 
           start = 2007, end = 2010)
